---
title: "Data Import and Cleaning"
author: "Lyuou Zhang"  
date: "3/27/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rnoaa)
library(rvest)
library(httr)
library(jsonlite)
library(lubridate)
library(ggridges)
library(sf)
```

## NOAA data

### Step 1. pull NOAA data from its API

I pulled all the stations in NY and used their IDs to pull all the data at these stations. `stations` is the dataset with all NY stations. `nydat` is the dataset that has the NY weather data

```{r eval = FALSE}
nystationids <- ghcnd_stations() %>% 
  filter(state == "NY") %>% 
  distinct(id)

# Pull the desired weather data for all of these stations
nydat <- meteo_pull_monitors(nystationids$id, 
                             date_min = "2007-01-01", 
                             date_max = "2017-12-31", 
                             var = c("PRCP", "SNOW", "SNWD", "TMAX", "TMIN"))
```

Note that this dataset does not have the county names. To find the county for each station, we used the latitude and longitude of each station and spatially join with the NYS county shapefile in GIS to get the county names.

```{r eval = FALSE}
ny_county <- stations %>% 
  filter(state == "NY") %>% 
  dplyr::select(id:name, -state) %>% 
  .[!duplicated(.[c('id', 'latitude', 'longitude', 'elevation', 'name')]),]

```

### Step 2: Join station IDs with counties, and with NOAA data

Some work behind the scene in QGIS...

After we figured out the county of each weather station, joined that back to the NOAA data in `nydat`

```{r eval = FALSE}
# import the ny county file joined with the county names from QGIS
ny_county_name <- read_csv('./data/ny_county_name.csv') %>% 
  janitor::clean_names() %>% 
  dplyr::select(-field_1) %>% 
  rename(county = name_1)
 

# join with nydata, and convert the temperature to degree celsius.
nydata <- nydat %>% 
  mutate(
    date = ymd(date)
  ) %>% 
  left_join(., ny_county_name, by = 'id') %>% 
  mutate(
    tmax = tmax/10,
    tmin = tmin/10
  )
 
```

Yayyy! So the final NOAA data will be `nydata`. Be sure to **remove all the `NA` values** before you use it.

## Birth data

This one is a lot easier.

```{r}
births <- read_delim('./data/NYS_0717_groupby_CYM_OEGestWeekly.txt', delim = '\t') %>% 
  janitor::clean_names() %>% 
  filter(is.na(notes)) %>% 
  filter(county_code != '36999') %>% 
  select(-notes)

```

## SES and population data

This part includes **education**, **household income** and **population** data import and cleaning. The data is from [American FactFinder](https://factfinder.census.gov/faces/nav/jsf/pages/index.xhtml).

### Education

*  I cleaned the data and only kept columns that are percent of population by each education level for people age **25 and over**.  
*  I also created two new variables:  
    *  `without_hs` is the percent of population without a high school degree  
    *  `college` is the percent of population with bachelors degree or higher

```{r}
education <- read_csv('./data/ACS_17_5YR_S1501_education/ACS_17_5YR_S1501_with_ann.csv') %>% 
  janitor::clean_names() %>% 
  dplyr::select(id2, contains('25_years_and_over'), -contains('total'), -contains('margin_of_error'), -contains('male')) %>% 
  dplyr::select(id2, percent_estimate_population_25_years_and_over_less_than_9th_grade:percent_estimate_population_25_years_and_over_graduate_or_professional_degree) %>% 
  rename(
    geoid = id2,
    less_than_9 = percent_estimate_population_25_years_and_over_less_than_9th_grade,
    grade_9_12 = percent_estimate_population_25_years_and_over_9th_to_12th_grade_no_diploma,
    high_school = percent_estimate_population_25_years_and_over_high_school_graduate_includes_equivalency,
    some_college = percent_estimate_population_25_years_and_over_some_college_no_degree,
    associates = percent_estimate_population_25_years_and_over_associates_degree,
    bachelors = percent_estimate_population_25_years_and_over_bachelors_degree,
    graduate_or_professional =  percent_estimate_population_25_years_and_over_graduate_or_professional_degree
  ) %>% 
  mutate(
    without_hs = less_than_9 + grade_9_12,
    college = bachelors + graduate_or_professional,
    geoid = as.character(geoid)
  )

```

### Household income

I only keep the median income for each household.

```{r}
income <- read_csv('./data/ACS_17_5YR_S1903_income/ACS_17_5YR_S1903_with_ann.csv') %>% 
  janitor::clean_names() %>% 
  dplyr::select(id2, median_income_dollars_estimate_households) %>% 
  rename(
    geoid = id2,
    med_income = median_income_dollars_estimate_households
  ) %>% 
  mutate(geoid = as.character(geoid))
  
```

### Population

Population should be in the model too (log(N)).

```{r warning = FALSE}
pop <- read_csv('./data/ACS_17_5YR_DP05_population/ACS_17_5YR_DP05_with_ann.csv') %>% 
  janitor::clean_names() %>% 
  dplyr::select(id2, estimate_sex_and_age_total_population) %>% 
  rename(
    geoid = id2,
    pop = estimate_sex_and_age_total_population
  ) %>% 
  mutate(geoid = as.character(geoid))

```

### Combine all the datasets

The final SES dataset is `ses.csv`

```{r}
ses <- education %>% 
  left_join(., income, by = 'geoid') %>% 
  left_join(., pop, by = 'geoid')

write.csv(ses, './data/ses.csv')
```




