---
title: "Mixed Models"
author: "Lyuou Zhang"
date: "3/28/2019"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(readr)
library(dplyr) 
library(lme4)
library(ICC) 
library(lubridate)
```

## Data Import

We are studying the impact of temperature variability on the number of preterm births, adjusted by SES status. We are using these data:  
  
*  Temperature data from NOAA: `nydata_month.csv`  
*  Birth data from CDC WONDER: `birth_2007_2010.csv`  
*  SES data from ACS: `ses.csv`  

Note that for analysis, I need to convert the birth data to binary: # of preterm birth and # of not preterm birth

```{r}
# read temp data
nydata_month <- read_csv('./data/nydata_month.csv') %>% 
  dplyr::select(-X1)

# read birth data
births <- read_csv('./data/birth_2007_2010.csv') %>% 
  dplyr::select(-X1, -year, -average_birth_weight, -month, -oe_gestational_age_weekly_code) %>% 
  mutate(
    county_code = as.character(county_code)
    ) %>% 
  spread(
    key = oe_gestational_age_weekly, value = births
  ) %>% 
  dplyr::select(-Unknown) %>% 
  mutate_all(funs(replace(., is.na(.), 0))) %>% 
  mutate(
    preterm = `24 weeks` + `25 weeks` + `26 weeks` + `27 weeks` + `28 weeks` + `29 weeks` + `30 weeks` + `31 weeks` + `32 weeks` + `33 weeks` + `34 weeks` + `35 weeks` + `36 weeks` + `37 weeks`,
    non_preterm = `38 weeks` + `39 weeks` + `40 weeks` + `41 weeks` + `42 weeks` + `43 weeks`
    ) %>% 
  dplyr::select(-(`24 weeks`:`43 weeks`)) %>% 
  mutate(
    county = str_replace(county, ' County, NY', '')
  )
  

# import ses data
ses <- read_csv('./data/ses.csv') %>% 
  dplyr::select(-X1) %>% 
  mutate(geoid = as.character(geoid))

```

## Between and within subject variablity

Here I use **ICC** to compare between and within subject variability.

**Criteria:**   
*  ICC > 0.75: Excellent reproducibility  
*  0.4 <= ICC < 0.75: Fair to good reproducibility   
*  ICC < 0.4: Poor reproducibility  

Note that the variables to use in the ICC should be the percentage of preterm births rather than the number of preterm births, because the latter depends on the total population of that county.

I will save the ICC results of outcome in `icc_birth`
```{r}
icc_birth <- births %>% 
  mutate(
    preterm_rate = preterm/(preterm + non_preterm)
  ) %>% 
  ICCest(county_code, preterm_rate, data = .)

```


Then the exposure variable, saved in `icc_temp`
```{r}
icc_temp <- births %>% 
  left_join(., nydata_month, by = 'county') %>% 
  ICCest(county_code, sd_temp, data = .)


```

*  ICC for Y: (`r icc_birth$LowerCI`, `r icc_birth$UpperCI`)
*  ICC for X: (`r icc_temp$LowerCI`, `r icc_temp$UpperCI`)

The ICC shows that there is much variability within subjects. 

## Mixed model



```{r}



```







